---
title: "Testing density options"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, echo = FALSE, eval = FALSE}
rmarkdown::render("density-playing.Rmd")
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.width = 10,
  fig.height = 8
)
```

```{r, packages}
library(dplyr)
library(devtools)
# library(pacea)
library(HDInterval)
library(kdensity)   # not going to end up using
library(spatstat.geom)   # for interval
library(BMS)        # for quantile.density
library(hdrcde)     # Hydman's, need to see about 0
load_all()
```

Wanted to resolve the zero issue, how to force density to 0 at 0, but with enough
resolution to get exact HDIs. Lots of ideas here but nothing that worked, which
could be the reason that we hadn't found any solutions for this problem
elsewhere.

So not resolvable, but for stock assessments that's okay, as it's the
recruitments that this issue came up, not relative biomass. And the projections
are done on the MCMC samples, so no real conservation issue there.

## Simulate from a lognormal then try different options

```{r simulate}
set.seed(42)
dat <- rlnorm(8000, meanlog = 1)
sort(dat)[1:10]            # Lowest 10 values
lowest <- min(dat)
lowest
```

After doing much of the later stuff (so the text won't compare to this calculation), realised I hadn't calculate the theoretical HDI.
```{r hditheoretical}
theory <- HDInterval::hdi(qlnorm, 0.95, meanlog = 1)
theory
plnorm(theory["lower"], meanlog = 1)
plnorm(theory["upper"], meanlog = 1)
```
Those seem close enough.

Just do the simple hdi with high n:
```{r hdiunlog}
HDInterval::hdi(density(dat, n = 1e05), meanlog = 1)
```
Not far off the theoretical. Then do with density cut off at 0 (looking into
more below)
```{r hdiunlogzero}
HDInterval::hdi(density(dat, n = 1e05, from = 0), meanlog = 1)
```
So it's pretty close to the theoretical. Maybe we just shouldn't worry too much
about the 0 issue after all (as Marie said on today's Zoom call).

```{r simulatedens}
dens <- density(dat)
dens
plot(dens)
dlnorm(0, meanlog = 1)     # pdf is 0 at x=0

```

Problem is that goes negative, when the 'true' lognormal should not. So set `from`:
```{r simulate2}
dens2 <- density(dat, from = 0)
dens2
plot(dens2)
dens2$x[1:10]
dens2$y[1:10]
HDInterval::hdi(dens2)
```

Problem is it does not fall to 0 at 0, as the resolution is not fine enough
there. Try increasing `n`:
```{r simulate3}
dens3 <- density(dat, from = 0, n = 1e05)
dens3
plot(dens3)
dens3$x[1:10]
dens3$y[1:10]
sum(dens3$x < lowest)      # Number below the lowest data point.
```

Problem still not resolved. Try whacking up `n` further:
```{r simulate4}
dens4 <- density(dat, from = 0, n = 1e07)
dens4
plot(dens4)
dens4$x[1:10]
dens4$y[1:10]
sum(dens4$x < lowest)      # Number below the lowest data point.
```

Still doesn't change. Marie said try reducing `adjust`:
```{r simulate5}
dens5 <- density(dat, from = 0, n = 1e07, adjust = 0.5)
dens5
plot(dens5)
dens5$x[1:10]
dens5$y[1:10]
sum(dens5$x < lowest)      # Number below the lowest data point.
```

Takes a long time to plot, more wiggly in the tail. Still doesn't reach 0.

Quick try of the `SJ` method mentioned in `?density`, putting `n` back
```{r simulate6}
dens6 <- density(dat, from = 0, bw = "SJ", n = 1e05)
dens6
plot(dens6)
dens6$x[1:10]
dens6$y[1:10]
sum(dens6$x < lowest)      # Number below the lowest data point.
```

Does get further down but more wiggly in tails. Marie hadn't thought algorithms
would help.

## Do density on a log scale then transform back

This is on the log scale:

```{r simulate7}
log_dat <- log(dat)
dens7 <- density(log_dat, n = 1e05)
dens7
plot(dens7)
dens7$x[1:10]
dens7$y[1:10]
```

Now transform back, still want to call it a density object. Don't think that
`plot()` uses the bandwidth, or assumes equally spaced `x` values (HDInterval
does not either)
```{r simulate7a}
dens7_unlog <- dens7                # Get the right format
dens7_unlog$x <- exp(dens7$x)
# need to transform y also:
dens7_unlog$y <- dens7$y / dens7_unlog$x
dens7_unlog

plot(dens7_unlog)
dens7_unlog$x[1:10]
dens7_unlog$y[1:10]
```

Looks good. Now check the intervals:
```{r simulate7b}
hdi_7 <- HDInterval::hdi(dens7_unlog)
hdi_7
plot(dens7_unlog, xlim = c(0, 40))
abline(h = attr(hdi_7, "height"))
abline(v = hdi_7["lower"])
abline(v = hdi_7["upper"])
```

Woo-hoo. That looks to work. Check with Marie, and do a few confirmation
calculations:


```{r simulate7c}
head(diff(dens7_unlog$x), 10)
diff(head(diff(dens7_unlog$x), 10))
tail(diff(dens7_unlog$x), 10)
diff(tail(diff(dens7_unlog$x), 10))
```
So the differences are still small, hence the smoothness of the plot.

So the `dens7_unlog` that worked did not give a warning about the interval being
split, presumably because it is smoothed out enough.

Check the integral of the HDI:

```{r simulate7d}
spatstat.geom::integral(dens7_unlog)
spatstat.geom::integral(dens7_unlog,
                        domain = c(as.numeric(hdi_7["lower"]),
                                   as.numeric(hdi_7["upper"])))
```

wtf, should be 0.95.

Marie realised the HDI is wrong because of the `cumsum` aspect in
`hdi.density()`. It takes a density argument and assumes the `x` values are
equally spaced, but they're not (as I fudged it). But we can scale them by the
widths between values to calculate the HDI and also check the integral (there's
a `weight` option in `spatstat.geom::integral.density()` that should presumably
result in the same answer; also that function doesn't seem to assume equally
spaced x).

```{r scaled}
dens7_unlog_scaled <- dens7_unlog     # To get the same format. Going to give
                                      # midpoints of x gaps and the density
                                      # being the mean height for each gap; so
                                      # one less value
scaled_dx <- diff(dens7_unlog$x)      # Differences between x values.
dens7_unlog_scaled$x <- dens7_unlog$x[-length(dens7_unlog$x)] + scaled_dx / 2  # Gives midpoints
dens7_unlog_scaled$y <- (dens7_unlog$y[-1] +
                         dens7_unlog$y[-length(dens7_unlog$y)]) / 2
# Gives scaled density that takes into account each x gap
plot(dens7_unlog_scaled, xlim = c(0, 30))
```

Now do the HDI on this, and also the interval:
```{r scaled2}
hdi_7_scaled <- HDInterval::hdi(dens7_unlog_scaled)
hdi_7_scaled
plot(dens7_unlog_scaled, xlim = c(0, 20))
abline(h = attr(hdi_7_scaled, "height"))
abline(v = hdi_7_scaled["lower"])
abline(v = hdi_7_scaled["upper"])
```

Numbers look very similar to `hdi_7`, which is slightly unexpected?

Now check the integrals:

```{r scaled3}
spatstat.geom::integral(dens7_unlog_scaled)
spatstat.geom::integral(dens7_unlog_scaled,
                        domain = c(as.numeric(hdi_7_scaled["lower"]),
                                   as.numeric(hdi_7_scaled["upper"])))
```

wtff - still not close enough. Maybe it's the integral calculation, especially
with all the trapezoid assumptions in the `integral.density()`.

What does it give for each tail:
```{r scaled4}
spatstat.geom::integral(dens7_unlog_scaled,
                        domain = c(0,
                                   as.numeric(hdi_7_scaled["lower"])))

spatstat.geom::integral(dens7_unlog_scaled,
                        domain = c(as.numeric(hdi_7_scaled["upper"]),
                                   Inf))
```
So it's the upper tail that seems (more) inaccurate.


Can check the integral on the density calculated on a regular grid (do a new one
here):
```{r scaled5}
dens_reg <- density(dat, n = 1e04)
spatstat.geom::integral(dens_reg)
spatstat.geom::integral(dens_reg,
                        domain = c(as.numeric(hdi_7_scaled["lower"]),
                                   as.numeric(hdi_7_scaled["upper"])))
spatstat.geom::integral(dens_reg,
                        domain = c(as.numeric(hdi_7_scaled["upper"]),
                                   Inf))

```

Nope.

So it could be the resolution of the grid around the upper end of HDI (kind of
what we first got into early on).
```{r scaled6}
hdi_7_scaled["lower"]
lower_ind <- findInterval(hdi_7_scaled["lower"],
                          dens_reg$x)   # approx correct, see create_intervals
                                        # if want to make exact
lower_ind
dens_reg$x[lower_ind + seq(-5, 5)]
dens_reg$y[lower_ind + seq(-5, 5)]


upper_ind <- findInterval(hdi_7_scaled["upper"],
                          dens_reg$x)   # approx correct, see above
dens_reg$x[upper_ind + seq(-5, 5)]
dens_reg$y[upper_ind + seq(-5, 5)]

cumsum(dens_reg$y[upper_ind])
```

## Maybe don't need such a high n when doing on the log scale

Repeat above but with `n` back to default, at least see how it changes calcs.

Copying it all then editing/deleting as needed:

```{r simulate10}
log_dat <- log(dat)
dens10 <- density(log_dat)
dens10
plot(dens10)
dens10$x[1:10]
dens10$y[1:10]
```

Now transform back, still want to call it a density object. Don't think that
`plot()` uses the bandwidth, or assumes equally spaced `x` values (HDInterval
does not either)
```{r simulate10a}
dens10_unlog <- dens10                # Get the right format
dens10_unlog$x <- exp(dens10$x)
# need to transform y also:
dens10_unlog$y <- dens10$y / dens10_unlog$x
dens10_unlog

plot(dens10_unlog)
dens10_unlog$x[1:10]
dens10_unlog$y[1:10]
```

Looks good. Now check the intervals:
```{r simulate10b}
hdi_10 <- HDInterval::hdi(dens10_unlog)
hdi_10
plot(dens10_unlog, xlim = c(0, 40))
abline(h = attr(hdi_10, "height"))
abline(v = hdi_10["lower"])
abline(v = hdi_10["upper"])
```

That works again. Do a few confirmation
calculations:

```{r simulate10c}
head(diff(dens10_unlog$x), 10)
diff(head(diff(dens10_unlog$x), 10))
tail(diff(dens10_unlog$x), 10)
diff(tail(diff(dens10_unlog$x), 10))
```
So the differences are still small, hence the smoothness of the plot.

Check the integral of the HDI:

```{r simulate10d}
spatstat.geom::integral(dens10_unlog)
spatstat.geom::integral(dens10_unlog,
                        domain = c(as.numeric(hdi_10["lower"]),
                                   as.numeric(hdi_10["upper"])))
```

wtf, should be 0.95. But still 0.81. Try the other ideas we had earlier:

```{r scaled10}
dens10_unlog_scaled <- dens10_unlog     # To get the same format. Going to give
                                      # midpoints of x gaps and the density
                                      # being the mean height for each gap; so
                                      # one less value
scaled_dx <- diff(dens10_unlog$x)      # Differences between x values.
dens10_unlog_scaled$x <- dens10_unlog$x[-length(dens10_unlog$x)] + scaled_dx / 2  # Gives midpoints
dens10_unlog_scaled$y <- (dens10_unlog$y[-1] +
                         dens10_unlog$y[-length(dens10_unlog$y)]) / 2
# Gives scaled density that takes into account each x gap
plot(dens10_unlog_scaled, xlim = c(0, 30))
```

Now do the HDI on this, and also the interval:
```{r scaled11}
hdi_10_scaled <- HDInterval::hdi(dens10_unlog_scaled)
hdi_10_scaled
plot(dens10_unlog_scaled, xlim = c(0, 20))
abline(h = attr(hdi_10_scaled, "height"))
abline(v = hdi_10_scaled["lower"])
abline(v = hdi_10_scaled["upper"])
```

Numbers look very similar to `hdi_10` and earlier ones I think; these values
aren't changing much.

Now check the integrals:

```{r scaled12}
spatstat.geom::integral(dens10_unlog_scaled)
spatstat.geom::integral(dens10_unlog_scaled,
                        domain = c(as.numeric(hdi_10_scaled["lower"]),
                                   as.numeric(hdi_10_scaled["upper"])))
```

wtff - still not close enough. Maybe it's the integral calculation, especially
with all the trapezoid assumptions in the `integral.density()`.

What does it give for each tail:
```{r scaled13}
spatstat.geom::integral(dens10_unlog_scaled,
                        domain = c(0,
                                   as.numeric(hdi_10_scaled["lower"])))

spatstat.geom::integral(dens10_unlog_scaled,
                        domain = c(as.numeric(hdi_10_scaled["upper"]),
                                   Inf))
```
So it's the upper tail that seems (more) inaccurate.


Can check the integral on the density calculated on a regular grid (do a new one
here):
```{r scaled14}
dens_reg <- density(dat, n = 1e04)
spatstat.geom::integral(dens_reg)
spatstat.geom::integral(dens_reg,
                        domain = c(as.numeric(hdi_10_scaled["lower"]),
                                   as.numeric(hdi_10_scaled["upper"])))
spatstat.geom::integral(dens_reg,
                        domain = c(as.numeric(hdi_10_scaled["upper"]),
                                   Inf))

```

Nope.

So it could be the resolution of the grid around the upper end of HDI (kind of
what we first got into early on).
```{r scaled15}
hdi_10_scaled["lower"]
lower_ind <- findInterval(hdi_10_scaled["lower"],
                          dens_reg$x)   # approx correct, see create_intervals
                                        # if want to make exact
lower_ind
dens_reg$x[lower_ind + seq(-5, 5)]
dens_reg$y[lower_ind + seq(-5, 5)]


upper_ind <- findInterval(hdi_10_scaled["upper"],
                          dens_reg$x)   # approx correct, see above
dens_reg$x[upper_ind + seq(-5, 5)]
dens_reg$y[upper_ind + seq(-5, 5)]

cumsum(dens_reg$y[upper_ind])
```



END of copying all

## Found hdrcde package

Try hdrcde package.
```{r hdrcde}
res_h <- hdr(dat, prob = 95)
res_h
```
So default goes below 0, as expected. Try the options, though the help says to
use $lambda < 1$ to force non-zero only, but not clear exactly what to set it at.
```{r hdrcde2}
res_h2 <- hdr(dat, prob = 95, lambda = 0)
res_h2

res_h3 <- hdr(dat, prob = 95, lambda = 0.25)
res_h3

res_h4 <- hdr(dat, prob = 95, lambda = 0.5)
res_h4

res_h5 <- hdr(dat, prob = 95, lambda = 0.9)
res_h5

```
Don't make too much sense to me.

What about using the density from earlier:
```{r hrdcde3}
dens7_unlog
res_h5 <- hdr(den = dens7_unlog, prob = 95)
res_h5

spatstat.geom::integral(dens7_unlog)   # as above
spatstat.geom::integral(dens7_unlog,
                        domain = c(as.numeric(res_h5$hdr[,1]),
                                   as.numeric(res_h5$hdr[,2])))
```
Again, not 0.95. Maybe the integral function is messing up. Find something else.



## Further ideas, but not as good

Calculate the interval on the linear scale, with high `n`, and no `from = 0`:
```{r simulate8}
dens8 <- density(dat, n = 1e05)
hdi_8 <- HDInterval::hdi(dens8)
hdi_8
plot(dens8, xlim = c(0, 40))
abline(h = attr(hdi_8, "height"))
abline(v = hdi_8["lower"])
abline(v = hdi_8["upper"])
```

The lines actually match up, but the lower end is unrealistically negative.

And with `from = 0` we get:
```{r simulate9}
dens9 <- density(dat, n = 1e05, from = 0)
hdi_9 <- HDInterval::hdi(dens9)
hdi_9
plot(dens9, xlim = c(0, 40))
abline(h = attr(hdi_9, "height"))
abline(v = hdi_9["lower"])
abline(v = hdi_9["upper"])
```
which will mess up the plotting, and I'm not sure integrates correctly.


## Try the kdensity package

Came across `kdensity` package (which does have a few dependencies, which would be nice to
avoid), basically includes a forced zero option to avoid boundary bias on the
half line. And it seems to be being updated.

Try it here, using their example

```{r kdensity}
kde <- kdensity(airquality$Wind, start = "gamma", kernel = "gamma")
plot(kde, main = "Wind speed (mph)")
lines(kde, plot_start = TRUE, col = "red")   # not clear on README what the
                                        # difference is, have to dig a bit
lines(density(airquality$Wind, adjust = 2), col = "blue")   # usual density,
                                        # goes below 0
rug(airquality$Wind)
```

But I don't think it will work with HDIinterval, as not of class `density`:
```{r kdensity2}
class(kde)
# HDInterval::hdi(kde)    # Gives Error, as no applicable method
```

Could check how the log transform idea matches with this though:
```{r kdelog}
dens_wind_log <- density(log(airquality$Wind), n= 1e05)
dens_wind_unlog <- dens_wind_log     # Get the right format
dens_wind_unlog$x <- exp(dens_wind_log$x)
dens_wind_unlog$y <- dens_wind_log$y / dens_wind_unlog$x

plot(kde, main = "Wind speed (mph)")
lines(kde, plot_start = TRUE, col = "red")
lines(dens_wind_unlog, col = "green")
```
Okay, not great, but very sparse data, way less than we'd have.

So worth looking at but doesn't look usable.
