---
title: "Analysis for Pacific Hake Numbers at Age"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis for Pacific Hake Numbers at Age}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("hake-natage.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

This analyses Pacific Hake assessment estimates of numbers-at-age, to produce
the age-1 result quoted in the manuscript, that the total reduction in uncertainty
over the time series when using HDIs instead of ETIs is >23 billion fish. This
report calculates values for all age classes.

Raw MCMC values are not included in
the package (but are available on request) since the file is 65 Mb, as
it contains 8,000 MCMC samples for each age 0 to 20 for 59 years. Thus, this
report will not run for users, but the code is all here and the rendered .html
is saved on GitHub.



```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)

redo_results <- FALSE   # If FALSE then load in previous values and results, since takes a
                        # while to do the calculations.
```

## Get the MCMC values and wrangle into the required format

```{r loaddata}
if(redo_results){
  model_path <- paste0(here::here(),
                       "/data-raw/hake-2024/02-version/01-base-models/01-base/")
  natage_mcmc <- readRDS(paste0(model_path, "01-base-natage.rds"))
  saveRDS(natage_mcmc, "natage_mcmc.rds")
} else {
  natage_mcmc <- readRDS("natage_mcmc.rds")
}

summary(natage_mcmc)
dim(natage_mcmc)
```

So 59 years and 8000 samples gives 472,000 rows. Done by iteration 1, then each
year, then iteration 2, etc. Ages are columns. And numbers of fish are in
millions here (as for Table 17, but will convert to billions below for
consistency in manuscript).

Make a list, with components that are each a tibble for a given age, in the same
format as `hake_recruitment_mcmc` (except no Virgin value), then we can use the same code:

```{r rec}
hake_recruitment_mcmc
```

So create the list:
```{r natagelist}
years <- unique(natage_mcmc$yr)
ages <- as.numeric(names(natage_mcmc[-c(1, 2)]))

if(redo_results){
  hake_natage_mcmc_list <- list()

  # Convert from millions to billions
  natage_mcmc_billions <- natage_mcmc
  index <- which(!names(natage_mcmc) %in% c("yr", "iter"))   # which columns are
                                                             # numbers to rescale
  natage_mcmc_billions[ , index]  <- natage_mcmc_billions[, index] / 1000

  for(i in ages){
    natage_this_age <- select(natage_mcmc_billions,
                              yr,
                              iter,
                              as.character(all_of(i)))

    natage_this_age_wide <- tidyr::pivot_wider(natage_this_age,
                                               names_from = yr,
                                               values_from = as.character(i))

    hake_natage_mcmc_list[[paste0("age", i)]] <- select(natage_this_age_wide,
                                                        -iter)
  }
  saveRDS(hake_natage_mcmc_list, "hake_natage_mcmc_list.rds")
} else {
  hake_natage_mcmc_list <- readRDS("hake_natage_mcmc_list.rds")
}
```

Each element of the list is a tibble in the same format as
`hake_recruitment_mcmc`, named as age0, age1, etc. up to age20.

## Do the analysis for each age in turn

Now do the analysis for each age. Resulting list has an element for each age,
which is itself a list object with the results for each year plus the summary in
`intervals_all`, which includes the median which we can use for checking (or use
each year's results for plotting).

```{r analysis}
if(redo_results){
  hake_natage_res_list <- list()

  for(ii in ages){
    hake_natage_res_list[[paste0("age", ii)]] <- create_intervals(hake_natage_mcmc_list[[paste0("age", ii)]])
  }
  saveRDS(hake_natage_res_list, "hake_natage_res_list.rds")
} else {
  hake_natage_res_list <- readRDS("hake_natage_res_list.rds")
}
```

Then check recruitment values agree with those calculated in Results vignette:
```{r checkage0}
# hake_natage_res_list[[paste0("age", i)]]
hake_natage_res_list[["age0"]]$intervals_all %>% filter(quantity == 2021) %>% a()
```

Last result matches those for recruitment (as they should) in 2021 already in
manuscript.

## Figure 2a equivalent for all ages

Can generate an equivalent to Figure 2a from these results, doing age-0 (to
confirm it matches Figure 2a), then age-10 as an example, and then we will show them all
after looking at the results:
```{r age0plot}
plot(hake_natage_res_list[["age0"]],
     xlab = "Year",
     ylab = "Age-0 fish (billions)")

plot(hake_natage_res_list[["age10"]],
     xlab = "Year",
     ylab = "Age-10 fish (billions)")
```

Now work out the total difference in interval widths across all years, for each
age class in turn. Have to exclude years that are uninformed by data, namely the
final two years for age-0 and the final year for age-1. Should
probably exclude some of the early years for the older fish, since they are
informed by limited data (if they are so old when the data starts, there won't
be many years of future age data about them), but not reporting results in the manuscript and so only do
that if needed.

```{r widthdiffage}
total_width_diff_by_age <- tibble(age = ages,
                                  total_width_diff = NA)

for(i in ages){
  if(i == 0){
    sum_this_age <- hake_natage_res_list[[paste0("age", i)]]$intervals_all %>%
      filter(!(quantity %in% c(2023, 2024))) %>%
      pull(width_diff) %>%
      sum()
  }

  if(i == 1){
    sum_this_age <- hake_natage_res_list[[paste0("age", i)]]$intervals_all %>%
      filter(!(quantity %in% c(2024))) %>%
      pull(width_diff) %>%
      sum()
  }

  if(i > 1){
    sum_this_age <- sum(hake_natage_res_list[[paste0("age", i)]]$intervals_all$width_diff)
  }

  total_width_diff_by_age[total_width_diff_by_age$age == i, "total_width_diff"] <- sum_this_age
}
# total_width_diff_by_age %>% a()
```

To see all the total differences for each age class:
```{r hakeagevalues}
knitr::kable(total_width_diff_by_age,
             digits = 1)
```

The age-0 value matches the recruitment one quoted in the text, and the table
provides the age-1 value. These differences remain substantial for many age
classes, diminishing in size as intuitively expected.

To plot the equivalent to Figure 2A for each age class, first with each y-axis
scaled independently:
```{r hakeeachage}
for(i in ages){
  plot(hake_natage_res_list[[paste0("age", i)]],
       xlab = "Year",
       ylab = paste0("Age-", i, " fish (billions)"))
}
```

Now with each y-axis the same scale, to emphasise the diminishing numbers of
fish as they age:
```{r hakeeachage2}
for(i in ages){
  plot(hake_natage_res_list[[paste0("age", i)]],
       xlab = "Year",
       ylab = paste0("Age-", i, " fish (billions)"),
       ylim = c(0, 35))
}
```
