---
title: "Vignette for hdiAnalysis"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for hdiAnalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("hdiAnalysis-vignette.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

We present what works, to be used as a template for other analyses, and then
change some code settings to show what did not work for our investigations.

## Figure 1

```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)
library(coda)            # Needed to test alternative code
```

ROUGH for now to remake figures

The hake Markov chain Monte Carlo (MCMC) recruitment and relative biomass
values are saved as data objects in the package, with `rec_2021` being the
vector of 8,000 MCMC values used for our Fig. 1.

```{r rec1}
summary(rec_2021)
length(rec_2021)
res_2021 <- create_intervals(rec_2021)
res_2021
res_2021$intervals %>% a()
```

Check the integral, using our `integrate_simpson()` function to evaluate an integral on
a density object:
```{r rec2}
integrate_simpsons(res_2021$density)
integrate_simpsons(res_2021$density,
                   domain = c(as.numeric(res_2021$intervals$eti_lower),
                              as.numeric(res_2021$intervals$eti_upper)))

integrate_simpsons(res_2021$density,
                   domain = c(as.numeric(res_2021$intervals$hdi_lower),
                              as.numeric(res_2021$intervals$hdi_upper)))
```

So the ETI approach is only a 94% interval, so it's actually narrower
than the true 95% ETI interval.

HDI is bang on, which is a good confirmation.

Plotting the resulting density function and intervals is simply done using our
custom plotting functions, with options for automatically adding intervals and lines. The ETI is:
```{r plot}
plot(res_2021,
     type = "eti",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of age-0 fish)",
     arrowhead_gap = 0.1)
```
and the HDI is
```{r plot2}
plot(res_2021,
     type = "hdi",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of age-0 fish)",
     arrowhead_gap = 0.1)
```
The `res_2021` object has our class `intervals_density`, and our custom plotting
function `plot.intervals_density()` automatically gets used on the object. See
`?plot.intervals_density()` for details and full options. Our complete Figure 1
is reproduced with the command `figure_1()`, which can be adapted by users if
desired for their own data.

## Figure 2A

The MCMC samples of hake recruitment for each year are saved in the package as
the tibble `hake_recruitment_mcmc`, with each column corresponding to a year
(with the first column representing the Virgin situation) and each of the 8,000
rows representing an MCMC sample:
```{r hake_mcmc}
hake_recruitment_mcmc
```

To create ETIs and HDIs for estimates of recruitment for each year, simply use
`create_intervals()` which, because `hake_recruitment_mcmc` has class
`data.frame`, uses our function `create_intervals.data.frame()` to automatically
calculate intervals for each column (year in this case, and we will exclude the
Virgin sample) of samples.
```{r hake_mcmc2}
res_all_years <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                -"Virgin"))
```
This gives a list object, with `res_all` being a list for which each element corresponds to a year and
gives the same results as above for a single year (and also with the `$name`
element corresponding the column of the input, a year in this example). So the
2021 results from above are:
```{r hake_mcmc3}
res_all_years$res_all[[56]]
```

For convenience, the intervals for all years are saved in a single tibble:
```{r hake_mcmc4}
res_all_years$intervals_all
```

To see the main values of interest, namely the ETIs and HDIs and the difference
between their widths:
```{r hake_mcmc5}
dplyr::select(res_all_years$intervals_all,
              quantity, eti_lower, eti_upper, hdi_lower, hdi_upper, width_diff) %>%
  a()
```

The sum of the differences over all years between the width of the ETI and width
of the HDI, but excluding 2023 and 2024 as these are not informed by data, is
```{r hake_mcmc6}
res_all_years$intervals_all %>%
  dplyr::filter(!quantity %in% c(2023, 2024)) %>%
  dplyr::pull(width_diff) %>%
  sum()
```
which is the source of the ``>25 billion fish`` statement reported in the main text.

To plot Fig. 2A, we have a custom plotting function
`plot.intervals_density_list()`, that gets called because `res_all_years` is
defined to have class `intervals_density_list`, so we can just use:
```{r hake_mcmc7}
plot(res_all_years)
```

The ETIs match those shown in Table 24 of the 2024 hake assessment (values
manually checked, code not run here)
```{r hake_mcmc8, eval = FALSE}
res_all_years$intervals_all %>%
  dplyr::select(eti_lower, eti_upper) %>%
  a() %>%
  round(3) * 1000
```

To plot all density plots and ETIs (might not be evaluated when editing as take a long time):
```{r hake_mcmc9, fig.height = 360, eval = FALSE}
plot(res_all_years,
     type = "eti",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(60, 1))  # 59 default
```

To plot all density plots and HDis (might not be evaluated when editing as take a long time):
```{r hake_mcmc9hdi, fig.height = 360, eval = FALSE}
plot(res_all_years,
     type = "hdi",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(60, 1))  # 59 default
```

And to see all the values:
```{r hake_mcmc9values}
# res_all_years$intervals_all %>% select(quantity, integral_full, integral_eti,
#                                       integral_hdi) %>%
#  print(n=60)
knitr::kable(res_all_years$intervals_all %>% select(quantity, integral_full,
                                                    integral_eti, integral_hdi),
             digits = 2)   # 3 shows that HDI are 95.0 except two 95.1, but this
                           # is easier to see.
```

## Figure 2B

TODO

The female spawning biomass is the estimated total biomass of all females in the
population that are mature (roughtly half of age-2 fish are considered mature).


## Absolute female spawning biomass

The above (TODO check) was for relative female spawning biomass. We now briefly look at
absolute female spawning biomass, in units of millions of tonnes. This is also
saved as a data object in the package (`hake_spawning_biomass_mcmc`). The values
for 2024 and 2023 are of interest and given in a bullet in the one-page summary of the hake assessment:

"The median estimate of female spawning biomass at the start of 2024 is 1,884,950 t
(with 95% credible interval from 853,207 to 4,828,382 t). This is an upward shift
from this assessment’s estimate for the 2023 female spawning biomass of 1,335,485 t
(with 95% credible interval from 652,495 to 3,224,819 t)."

So those credible intervals are ETIs. How would HDIs compare?

```{r sb1}
res_all_years_spawning <-
  create_intervals(dplyr::select(hake_spawning_biomass_mcmc,
                                 -"Virgin"))

res_all_years_spawning$intervals_all %>%
  dplyr::filter(quantity %in% c(2023, 2024)) %>%
  dplyr::select(quantity, eti_lower, eti_upper, hdi_lower, hdi_upper, width_diff) %>%
  a()
```
The ETIs agree with the above quote (except for 2023 the upper value is 1
tonne different). The credible interval using HDIs for 2024 is then
`r f(dplyr::filter(res_all_years_spawning$intervals_all, quantity ==
                                                       2024)$hdi_lower * 1e06)`
to
`r f(dplyr::filter(res_all_years_spawning$intervals_all, quantity ==
                                                       2024)$hdi_upper * 1e06)` t,
which is shifted downwards from the ETI of
`r f(dplyr::filter(res_all_years_spawning$intervals_all, quantity ==
                                                       2024)$eti_lower * 1e06)`
to
`r f(dplyr::filter(res_all_years_spawning$intervals_all, quantity ==
                                                       2024)$eti_upper * 1e06)` t.
The reduction in uncertainty is
`r f(dplyr::filter(res_all_years_spawning$intervals_all, quantity ==
                                                       2024)$width_diff * 1e06)` t,
which is larger than the total catch (of males and females, not just spawning
females) in the latest year of fishing (2023), and so is certainly not trivial.

The ETI and HDI plots are
```{r,  sb2, fig.height = 12}
par(mfrow = c(2, 1))
plot(res_all_years_spawning$res_all[[which(res_all_years_spawning$intervals_all$quantity == 2024)]],
     xlim = c(0, 8),
     ylim = c(0, 0.6),
     interval_arrows = TRUE,
     y_arrow = 0.57,
     xlab = "Female spawning biomass in 2024 (millions of tonnes)",
     type = "eti")

plot(res_all_years_spawning$res_all[[which(res_all_years_spawning$intervals_all$quantity == 2024)]],
     xlim = c(0, 8),
     ylim = c(0, 0.6),
     interval_arrows = TRUE,
     y_arrow = 0.57,
     xlab = "Female spawning biomass in 2024 (millions of tonnes)",
     type = "hdi")
```


## TODO

HERE - think some of this can go.

D First save the `allow_hdi_zero` as an output in create_intervals.
  have been thinking about this slightly wrong. min data equals hdi lower
  because we've restricted that, so then it's a one-sided tail for the HDI,
  which we should force to be the 5%. So if HDI starts at the min add a flag for
  this to help with plotting. Below the HDI is not a thing because we've
  restricted it, so just need to think differently for plotting.
D also save integral over (-Inf,Inf) and also over ETI and HDI.
- HERE it's drawing red lower for 2024 but shouldn't.
D if integral is not 1 then could rescale it; in example above that made the HDI
  contain 95%. Now done in create intervals.
- Then need to adapt plotting function to not add the 0,0 in if have forced the
  HDI to be above 0. Use that in plot to draw grey line, and to not add in 0,0.
- for text, first do we care about allowing zero (might not always),

```{r plotttt}
res_2024 <- create_intervals(select(hake_recruitment_mcmc, "2024") %>% pull())
plot(res_2024, xlim= c (0, 20))
```

Try changing adjust, changes wiggliness, and allow zero:
```{r adjust}
res_2024_allow_hdi_zero_adjust_0.1 <-
  create_intervals(select(hake_recruitment_mcmc, "2024") %>% pull(), adjust =
                                                                       0.1,
                   allow_hdi_zero = TRUE)
plot(res_2024_allow_hdi_zero_adjust_0.1, xlim = c(0, 1))

# Try 0.01:
res_2024_allow_hdi_zero_adjust_0.01 <-
  create_intervals(select(hake_recruitment_mcmc, "2024") %>% pull(), adjust =
                                                                       0.01,
                   allow_hdi_zero = TRUE)
plot(res_2024_allow_hdi_zero_adjust_0.01, xlim = c(0, 1))
plot(res_2024_allow_hdi_zero_adjust_0.01, xlim = c(0, 40))

```
Tweaking adjust (to 0.1 and the above 0.01) does not seem help, and just makes
the distribution really wiggly. Expect it makes the integrals strange - do the
above idea first about always calculating the integral, and the intergral over
the HDI.


# Investigating the issues we came across

## Issue 1 -- trying the default HDInterval options that is not based on density kernel

Explain (see Extended Methods).

```{r nodensity}
res_2021_no_density <- create_intervals(rec_2021,
                                        density = FALSE)
res_2021_no_density
res_2021_no_density$intervals %>% a()
```
Our code just uses the `HDInterval::hdi()` method, as confirmed here:
```{r nodensitycheck}
expect_equal(c(res_2021_no_density$intervals$hdi_lower,
               res_2021_no_density$intervals$hdi_upper),
             as.numeric(HDInterval::hdi(rec_2021)))
```


The density was calculated here but only for plotting purposes, and is the same
as above:
```{r nodensity1a}
expect_equal(res_2021_no_density$dens, res_2021$dens)
```

Also there is no `height` value because that was not used when calculating the
HDI.

The equivalent plot to our Fig. 1B for the calculated HDI is
```{r no_densityplot}
plot(res_2021_no_density,
     xlim = c(0,40),
     xlab = "Recruitment (billions of age-0 fish)")
```
But, as can be seen, the heights of the two endpoints of the HDI are not the
same, with the right-hand one being lower.

Furthermore, the integral over the
HDI is only 94.5%, not the theoretical 95%:
```{r nodensity4}
integrate_simpsons(res_2021_no_density$density,
                   domain = c(as.numeric(res_2021_no_density$intervals$hdi_lower),
                              as.numeric(res_2021_no_density$intervals$hdi_upper)))
```
This is what led us to investigate the other options.

The values of the HDI calculated using density and not using density are
different:
```{r nodensity5}
res_2021$intervals %>% dplyr::select(hdi_lower, hdi_upper)
res_2021_no_density$intervals %>% dplyr::select(hdi_lower, hdi_upper) %>% a()
```

Also try `coda::HPDinterval()`, as recommended in `emdbook::tcredint()` from Ben
Bolker's book.
```{r coda}
res_2021_coda <- coda::HPDinterval(coda::as.mcmc(rec_2021))
res_2021_coda
# Compared to:
res_2021$intervals %>% dplyr::select(hdi_lower, hdi_upper)
```
which are also the same as ours using `HDInterval::hdi()`:
```{r codacheck}
expect_equal(as.numeric(res_2021_coda),
             as.numeric(HDInterval::hdi(rec_2021)))
```

## Issue 2 - default kernel density approach did not quite work

The default of 512 equally spaced points (`n`) at which to compute the kernal density
estimate in the `density()` function, did not yield satisfactory
HDIs. Setting the number of points to 512 here shows this, giving a warning
(from our `create_intervals()` function  that the densities at the HDI lower and
upper values are not within a tolerance of 1% of each other, and the difference
can be seen in the plot:
```{r n512}
res_2021_n_512 <- create_intervals(rec_2021,
                                   n = 512)
plot(res_2021_n_512,
     xlim = c(0, 40))
```

We found that increasing `n` to 100,000 gave satisfactory results. Note that
increasing `n` does not increase the wiggliness of the kernel density estimate,
just the number of points at which it is calculated.

However, the integral of the density across this HDI with 512 points was the
correct 95%:
```{r n512integral}
integrate_simpsons(res_2021_n_512$density,
                   domain = c(as.numeric(res_2021_n_512$intervals$hdi_lower),
                              as.numeric(res_2021_n_512$intervals$hdi_upper)))
```

We also experimented with other `density()` settings, such as `adjust` and `bw`,
but the defaults were satisfactory.

## Issue 3 - zeros TODO

Have to figure out the options to use now, as the defaults changed.

After incorporating Issues 1 and 2,

We found that stil

With no `from` specified, we can get negative lower bounds of the HDI for recruitment:
```{r zeroissue}
res_all_years_from_null <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                          -c("Virgin", "1975")),
                                            from = NULL,
                                            allow_hdi_zero = TRUE)

# TODO figure out the 1975 error, to do with i_hdi_lower, think I should not
# have the ==
# Testing as get error
#res_all_years_from_null <- create_intervals(dplyr::select(hake_recruitment_mcmc,
#                                                          "1975"),
#                                            from = NULL,
#                                           allow_hdi_zero = TRUE)

res_all_years_from_null$intervals_all
sum(res_all_years_from_null$intervals_all$hdi_lower < 0)
```

Then, setting `from = 0` in `density()` prevents this, yet leads to the lower
bound of the HDI often being 0:

```{r checknonzero}
res_all_years_allow_hdi_zero <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                               -"Virgin"),
                                                 allow_hdi_zero = TRUE)

filter(res_all_years_allow_hdi_zero$intervals_all,
       hdi_lower == 0)

years_zero <- filter(res_all_years_allow_hdi_zero$intervals_all,
                     hdi_lower == 0)$quantity

# Check the zero years remaining calculationsa are the same
expect_equal(filter(res_all_years$intervals_all,
                    quantity %in% years_zero) %>%
             select(quantity, median, eti_lower, eti_upper),
             filter(res_all_years_allow_hdi_zero$intervals_all,
                    quantity %in% years_zero) %>%
             select(quantity, median, eti_lower, eti_upper))
```

But


## Investigate warnings regarding split intervals

The function `HDInterval::hdi()` gives a warning "The HDI is discontinuous but
allowSplit = FALSE; the result is a valid CrI but not HDI." The `intervals`
output from `create_intervals()` has a column `warning` that is `TRUE` if this
occurs; for example
```{r warning1}
res_2024$intervals %>% a()
```

This can be investigated using the plotting function, to check that there is
just a small region outside of the HDI that does in fact have a higher density
than some values inside the HDI, which is due to the density approximation being
slightly wiggly. This occurs for `res_2024` calculated above, and setting
`show_discontinuity = TRUE` plots points showing the region in question. These
can barely be seen in the default plot, so we then hone in on the region in
question, seeing that, indeed, it is only a small region and not something to
get overly concerned about.
```{r, warning2, fig.height = 12}
par(mfrow = c(2,1))
plot(res_2024, show_discontinuity = TRUE)

plot(res_2024,
     show_discontinuity = TRUE,
     xlim = c(0, 25),
     ylim = c(0, 0.1))
```

For the recruitments for all years, the following have the warning during the
calculation:
```{r warning3}
years_with_warning <- dplyr::filter(res_all_years$intervals_all,
                                    warning == TRUE)$quantity
years_with_warning
expect_equal(years_with_warning,
             c(1975, 2023, 2024))   # Change some of the following if get an error
```

The plots for 1975 and 2023 show similar unconcerning behaviour to that for
2024:
```{r warning4, fig.height = 12}
par(mfrow = c(2, 1))

plot(res_all_years$res_all[[which(res_all_years$intervals_all$quantity == 1975)]],
     show_discontinuity = TRUE)

plot(res_all_years$res_all[[which(res_all_years$intervals_all$quantity == 2023)]],
     show_discontinuity = TRUE,
     xlim = c(0, 25),
     ylim = c(0, 0.1))
```

Note that the 2023 and 2024 recruitments have no data to inform them and so
are essentially samples from the prior distribution used in the assessment.

So we recommend checking the plots to see that any such behaviour is
acceptable. The "HDI is discontinuous ..." warning does not get printed, but if
another warning occurs in the `HDInterval::hdi()` calculation then that will be
shown, along with "New type of warning in create_intervals().", which should be
likely be investigated.




TODO still

Deal with warnings, which are mine in `create_intervals()`. Just do 1966

NOT SURE IF THIS IS NEEDED, maybe as another example of warnings. Maybe plot
anything from earlier that had a warning. TODO

```{r warn}
res_2024_allow_hdi_zero <- create_intervals(select(hake_recruitment_mcmc, "2024") %>% pull(), allow_hdi_zero = TRUE)

res_2024_allow_hdi_zero$intervals %>% a()

res_2024 <- create_intervals(select(hake_recruitment_mcmc, "2024") %>% pull())

res_2024$intervals %>% a()

res_2024_allow_neg <- create_intervals(select(hake_recruitment_mcmc, "2024") %>%
                                         pull(),
                                       from = -2)

res_2024_allow_neg$intervals %>% a()
# HDI goes -ve, but allowing it to.

plot(res_2024_allow_neg, xlim = c(-2, 20))

# This should give the same, except allow_hdi_zero value. Yes it is, just
# commenting as it gives error for that one change
res_2024_allow_neg_allow_zero <- create_intervals(select(hake_recruitment_mcmc, "2024") %>%
                                                  pull(),
                                                  from = -2,
                                                  allow_hdi_zero = TRUE)

# expect_equal(res_2024_allow_neg$intervals,
#              res_2024_allow_neg_allow_zero$intervals)


# Prob don't need all this. Tidy up.
res_1966 <- create_intervals(select(hake_recruitment_mcmc, "1966") %>% pull())

# Might not have redone all these
expect_equal(min(select(hake_recruitment_mcmc, "2024")),
             as.numeric(res_2024$intervals$"hdi_lower"))
```



This is old and may not be needed any more as defaults changed:

The equivalent table for HDIs would instead be TODO adapt this to show a few
years, highlighting the zero issue
```{r hakemcmc8}
res_all_years$intervals_all %>%
  dplyr::select(hdi_lower, hdi_upper) %>%
  a() %>%
  round(3) * 1000
```


## Quick long chain, but easiest to just re-run vignette with long chain results used instead

Somewhere, check the longer chain, not sure if including in package (could do if
needed, or just save these calculations somewhere else and state we did them -
would really be redoing the vignette with longer chain. Do quick calcs here
while writing appendix
```{r longchain}
rec_2021_long <- dplyr::pull(hake_recruitment_mcmc_14_long,
                             "2021")

res_2021_long <- create_intervals(rec_2021_long)

res_2021_long_no_density <- create_intervals(rec_2021_long,
                                             density = FALSE)

res_2021_long$intervals %>% a()
res_2021_long_no_density$intervals %>% a()
# NOTE values change anyway, including median, so gets hard to compare with
# original
```

Do the plot:
```{r res2021longplot}
plot(res_2021_long_no_density,
     type = "hdi",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of age-0 fish)",
     arrowhead_gap = 0.1)
```

### Examples from HDInterval

Replicating the examples from the help file `?HDInterval::hdi` using our
functions to then easily make plots:
```{r hdinterval}
set.seed(42)
# for a vector:
tst <- rgamma(1e5, 2.5, 2)
HDInterval::hdi(tst)
HDInterval::hdi(tst, credMass=0.8)

# Use our code, and check values agree, and then plot:
res_tst_no_dens <- create_intervals(tst, density = FALSE, credibility = 0.8)
expect_equal(as.numeric(HDInterval::hdi(tst, credMass=0.8)),
             c(res_tst_no_dens$intervals$hdi_lower,
               res_tst_no_dens$intervals$hdi_upper))
plot(res_tst_no_dens)
```

```{r hdinterval2}
# For comparison, the symmetrical 80% CrI:
quantile(tst, c(0.1,0.9))

# for a density:
dens <- density(tst)
HDInterval::hdi(dens, credMass=0.8)
res_tst_dens <- create_intervals(tst, n = 512, credibility = 0.8)
res_tst_dens$intervals %>% a()
# The values don't exactly match because we have from = 0 as the default, and it
# is hard to then specific the default for density.
# expect_equal(as.numeric(hdi(dens, credMass=0.8)),
#              c(res_tst_dens$intervals$hdi_lower,
#                res_tst_dens$intervals$hdi_upper))
plot(res_tst_dens, xlim = c(0,4))
```

With our defaults for `n`, the uneveness is resolved:
```{r, hdinterval4}
res_tst_dens_default_n <- create_intervals(tst, credibility = 0.8)
res_tst_dens_default_n$intervals %>% a()
plot(res_tst_dens_default_n, xlim = c(0,4))

```
