---
title: "Analysis for Pacific Hake Numbers at Age"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis for Pacific Hake Numbers at Age}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("hdiAnalysis-vignette-hake-natage.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

Copying the current `hdiAnalysis-vignette-herring-sog.Rmd` (which cut out the
explanatory stuff) to look at numbers of
age-1 and age-2 fish for hake. TODO maybe move to report, rather than vignette,
then include what we endup reporoting as vignette (as then we can just save age1
or age2 full MCMC, saving all over all years is too much).

## Figure 1 equivalent

```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)
library(spatstat.geom)   # Needed for checking some integrals
library(coda)            # Needed to test alternative code

redo_results <- FALSE   # If FALSE then load in previous values and results, since takes a
                        # while to do the calculations.
```

May want to functionalise a little like `create_hake_forecast_mcmc()`.

TODO The hake numbers-at-age full MCMC results have been extracted and are saved
and then wrangled here, not saved in the package.

```{r loaddata}
if(redo_results){
  model_path <- paste0(here::here(),
                       "/data-raw/hake-2024/02-version/01-base-models/01-base/")
  natage_mcmc <- readRDS(paste0(model_path, "01-base-natage.rds"))
  saveRDS(natage_mcmc, "natage_mcmc.rds")
} else {
  natage_mcmc <- readRDS("natage_mcmc.rds")
}

summary(natage_mcmc)
dim(natage_mcmc)
```

So 59 years and 8000 samples gives 472,000 rows. Done by iteration 1, then each
year, then iteration 2, etc. Ages are columns. And numbers of fish are in
millions here (as for Table 17, but will convert to billions below for
consistency in manuscript).

Make a list, with components that are each a tibble for a given age, in the same
format as `hake_recruitment_mcmc` (except no Virgin value), then we can use the same code:

```{r rec}
hake_recruitment_mcmc
```

So create the list:
```{r natagelist}
years <- unique(natage_mcmc$yr)
ages <- as.numeric(names(natage_mcmc[-c(1, 2)]))

if(redo_results){
  hake_natage_mcmc_list <- list()

  # Convert from millions to billions
  natage_mcmc_billions <- natage_mcmc
  index <- which(!names(natage_mcmc) %in% c("yr", "iter"))   # which columns are
                                                             # numbers to rescale
  natage_mcmc_billions[ , index]  <- natage_mcmc_billions[, index] / 1000

  for(i in ages){
    natage_this_age <- select(natage_mcmc_billions,
                              yr,
                              iter,
                              as.character(all_of(i)))

    natage_this_age_wide <- tidyr::pivot_wider(natage_this_age,
                                               names_from = yr,
                                               values_from = as.character(i))

    hake_natage_mcmc_list[[paste0("age", i)]] <- select(natage_this_age_wide,
                                                        -iter)
  }
  saveRDS(hake_natage_mcmc_list, "hake_natage_mcmc_list.rds")
} else {
  hake_natage_mcmc_list <- readRDS("hake_natage_mcmc_list.rds")
}
```

Each element of the list is a tibble in the same format as
`hake_recruitment_mcmc`, named as age0, age1, etc. up to age20.

Now do the analysis for each age. Resulting list has an element for each age,
which is itself a list object with the results for each year plus the summary in
`intervals_all`, which includes the median which we can use for checking (or use
each year's results for plotting).

```{r analysis}
if(redo_results){
  hake_natage_res_list <- list()

  for(ii in ages){
    hake_natage_res_list[[paste0("age", ii)]] <- create_intervals(hake_natage_mcmc_list[[paste0("age", ii)]])
  }
  saveRDS(hake_natage_res_list, "hake_natage_res_list.rds")
} else {
  hake_natage_res_list <- readRDS("hake_natage_res_list.rds")
}
```

Then check recruitment:
```{r checkage0}
# hake_natage_res_list[[paste0("age", i)]]
hake_natage_res_list[["age0"]]$intervals_all %>% filter(quantity == 2021) %>% a()
```

Last result matches those for recruitment (as they should) in 2021 already in
manuscript. Should be able to generate Figure 2a from these results, then look
at example ages one at a time.

So age-0 over time, then age-10 and age-20
```{r age0plot}
plot(hake_natage_res_list[["age0"]])
plot(hake_natage_res_list[["age10"]])
plot(hake_natage_res_list[["age20"]])
```

See the large uncertainty in age-20 at the start of uncertainty, since pretty
uninformed by data; makes sense.

Now work out the total difference in interval widths across all years, for each
age class in turn (can't really just add all of those up as that's double
counting). Have to exclude years that are uninformed by data, which would
include early years for older fish, but we likely won't report those anyway so
don't worry.

```{r widthdiffage}
total_width_diff_by_age <- tibble(age = ages,
                                  total_width_diff = NA)

for(i in ages){
  if(i == 0){
    sum_this_age <- hake_natage_res_list[[paste0("age", i)]]$intervals_all %>%
      filter(!(quantity %in% c(2023, 2024))) %>%
      pull(width_diff) %>%
      sum()
  }

  if(i == 1){
    sum_this_age <- hake_natage_res_list[[paste0("age", i)]]$intervals_all %>%
      filter(!(quantity %in% c(2024))) %>%
      pull(width_diff) %>%
      sum()
  }

  if(i > 1){
    sum_this_age <- sum(hake_natage_res_list[[paste0("age", i)]]$intervals_all$width_diff)
  }

  total_width_diff_by_age[total_width_diff_by_age$age == i, "total_width_diff"] <- sum_this_age
}
total_width_diff_by_age %>% a()
```

And to see all the values:
```{r hakeagevalues}
knitr::kable(total_width_diff_by_age,
             digits = 1)
```

To plot equivalent to Figure 2A for each age class, first with each y-axis
scaled independntly:
```{r hakeeachage}
for(i in ages){
  plot(hake_natage_res_list[[paste0("age", i)]],
       xlab = "Year",
       ylab = paste0("Age-", i, " fish (billions)"))
}
```

Now with each y-axis the same scale, to emphasise the diminishing numbers of
fish as they age:
```{r hakeeachage2}
for(i in ages){
  plot(hake_natage_res_list[[paste0("age", i)]],
       xlab = "Year",
       ylab = paste0("Age-", i, " fish (billions)"),
       ylim = c(0, 35))
}
```
