---
title: "Extra results and examples of code functionality"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extra results and examples of code functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("results-extra.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

This vignette follows on from the `results.Rmd` vignette, including:

- plotting of all density functions when multiple years of data are analysed
(specifically all years of recruitment shown in Figure 2A).
- calculation on a longer MCMC chain that used for the hake assessment.
- exporation of additional options for calculating HDIs, based on the kernel
density estimate of the MCMC sample.

Set up:
```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)
```

## MCMC samples for multiple years -- recruitment for Figure 2A, plotting all the distributions

For the multiple years of MCMC samples of hake recruitment, we calculate the
results as in `results.Rmd` (see that for more details) and plot the full
distributions with ETI and HDI for each year.
```{r hake_mcmc2}
res_all_years <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                -"Virgin"),
                                  density = FALSE)
```

The default plot (which uses `plot.intervals_density_list()`) is that shown in Figure 2A:
```{r hake_mcmc7}
plot(res_all_years,
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
```

To plot all density plots and ETIs for each year (equivalent to Figure 1A), the `type` argument is used
(and we have changed the `fig.height` for the chunk, and specified `mfrow` as an
argument, and fixed `xlim` so that it is the same for each plot). The year is
automatically shown in each figure.
```{r hake_mcmc9, fig.height = 80}
plot(res_all_years,
     type = "eti",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(30, 2))  # 59 default
```

For the equvalent figures for HDIs, just change the `type` argument:
```{r hake_mcmc9hdi, fig.height = 80}
plot(res_all_years,
     type = "hdi",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(30, 2))  # 59 default
```

## Longer chain of 32,000 MCMC samples for hake recruitment in 2021

We also analysed an experimental longer chain of 32,000 MCMC samples (that took
several days to run), and results are similar to the standard chain of 8,000
samples. Code is not run here (since the data object is large and so not included
in the package) but results are presented below.

```{r longchain, echo = FALSE, eval = FALSE}
vec_long <- dplyr::pull(hake_recruitment_mcmc_14_long,
                        "2021")

res_vec_long <- create_intervals(vec_long,
                                  density = FALSE)
summary_table(res_vec_long)

plot(res_vec_long,
     type = "hdi",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of age-0 fish)",
     arrowhead_gap = 0.1)
# Density kernel is a bit smoother
```

| Quantity| Value | Interval width |
| :-------| -----:| -----:|
| Median  | 10.20 | -- |
| ETI     | 4.05 -- 29.53 | 25.49 |
| HDI     | 2.89 -- 25.23 | 22.34 |
| Range a | 2.27 -- 4.05 | 1.78 |
| Range b | 16.02 -- 29.53 | 13.52 |

The slight differences between the HDI for the two chains are small
compared to the difference to the ETI (for either chain). The ETIs change
slightly between changes due to randomness between them, and the median even
changes very slightly.
