---
title: "Simpson's simulation vignette"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simpson's simulation vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("simpsons-simulation.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

Doing some simulations to check some things regarding Simpson's Rule code. May
not be needed in final version, or move to report.

```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)
```

Simulated values are saved in the `sim` list object, so just use those.

```{r ints}
mean_log <- sim$mean_log
sd_log <- sim$sd_log
n <- length(sim$values)
mean_log
sd_log
n
```

Do repeated simulations with the same parameters for a lognormal, then
maybe need to do for multiple different parameters.

```{r simulations}
num_sims <- 1000
simulations <- list()
n_for_density <- formals(create_intervals.numeric)$n
ints_simulations <- data.frame(sim_num = 0,   # 0 for theoretical, then easy to plot
                          full = plnorm(Inf, mean_log, sd_log),
                          int_to_mean = plnorm(exp(mean_log), mean_log, sd_log),
                          int_to_100 = plnorm(100, mean_log, sd_log),   # 100 is a bit arbitrary
                          # Doing normalised kernels below, the theoretical
                          # values are the same here though:
                          full_norm = plnorm(Inf, mean_log, sd_log),
                          int_to_mean_norm = plnorm(exp(mean_log), mean_log, sd_log),
                          int_to_100_norm = plnorm(100, mean_log, sd_log),
                          full_spatstat = 1,
                          int_to_mean_spatstat = plnorm(exp(mean_log), mean_log, sd_log),
                          int_to_100_spatstat = plnorm(100, mean_log, sd_log))      # For simulations will calc
                                        # using spatstat
                          # Change below if this changes.

for(i in 1:num_sims){
  set.seed(i)
  this_sim <- rlnorm(n,
                     mean_log,
                     sd_log)

  this_dens <- density(this_sim,
                              n = n_for_density)

  this_int_full <- integrate_simpsons(this_dens)

  this_int_to_mean <- integrate_simpsons(this_dens,
                                         c(min(this_dens$x),
                                           exp(mean_log)))
  # Just to 100 (kind of arbitrarily)
  this_int_to_100 <- integrate_simpsons(this_dens,
                                         c(min(this_dens$x),
                                           100))

  # Then normalise the density and do the calculations, see if they are closer
  # to theoretical
  this_dens_norm <- this_dens
  this_dens_norm$y <- this_dens_norm$y/this_int_full

  this_int_full_norm <- integrate_simpsons(this_dens_norm)

  this_int_to_mean_norm <- integrate_simpsons(this_dens_norm,
                                              c(min(this_dens_norm$x),
                                                exp(mean_log)))
  # Just to 100 (kind of arbitrarily)
  this_int_to_100_norm <- integrate_simpsons(this_dens_norm,
                                             c(min(this_dens_norm$x),
                                               100))

  # Calculate integrals using spatstat::integral() on unnormalised, check how they compare to mine:
  this_int_full_spatstat <- spatstat.geom::integral(this_dens)

  this_int_to_mean_spatstat <- spatstat.geom::integral(this_dens,
                                         domain = c(min(this_dens$x),
                                                    exp(mean_log)))
  # Just to 100 (kind of arbitrarily)
  this_int_to_100_spatstat <- spatstat.geom::integral(this_dens,
                                         domain = c(min(this_dens$x),
                                                    100))


#  simulations[[i]] <- list("values" = this_sim,
#                           "ints_sim" = this_int_sim)

  ints_simulations <- rbind(ints_simulations,
                            c(i,
                              this_int_full,
                              this_int_to_mean,
                              this_int_to_100,
                              this_int_full_norm,
                              this_int_to_mean_norm,
                              this_int_to_100_norm,
                              this_int_full_spatstat,
                              this_int_to_mean_spatstat,
                              this_int_to_100_spatstat))

  if(i == 1){
    plot(this_dens,
         main = "Kernel density estimate for first iteration")
  }

}
ints_simulations <- tibble::as_tibble(ints_simulations)
ints_simulations

# These are in the plots
# sum(ints_simulations$full < pull(ints_simulations[1, "full"]))
# sum(ints_simulations$int_to_mean < pull(ints_simulations[1, "int_to_mean"]))
# sum(ints_simulations$int_to_100 < pull(ints_simulations[1, "int_to_100"]))
```

### Plot integration results

Now to plot the values, ordering the results in each plot:
```{r plotints}
# Full integral
plot(sort(ints_simulations$full),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Full integral",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for full integral is ",
                   sum(ints_simulations$full < pull(ints_simulations[1, "full"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(full),
       col = "red")

# Integral to mean (which is the mode, so not overly surprising okay if not perfect)
plot(sort(ints_simulations$int_to_mean),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to mean",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to mean is ",
                   sum(ints_simulations$int_to_mean < pull(ints_simulations[1, "int_to_mean"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_mean),
       col = "red")

# Integral to 100 (slighlty arbitrary)
plot(sort(ints_simulations$int_to_100),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to 100",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to 100 is ",
                   sum(ints_simulations$int_to_100 < pull(ints_simulations[1, "int_to_100"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_100),
       col = "red")

```

### Plot integration results using normalised density kernels

```{r plotintsnorm}
# Full integral
plot(sort(ints_simulations$full_norm),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Full integral",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for full integral is ",
                   sum(ints_simulations$full_norm < pull(ints_simulations[1, "full_norm"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(full_norm),
       col = "red")

# Integral to mean (which is the mode, so not overly surprising okay if not perfect)
plot(sort(ints_simulations$int_to_mean_norm),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to mean",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to mean is ",
                   sum(ints_simulations$int_to_mean_norm < pull(ints_simulations[1, "int_to_mean_norm"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_mean_norm),
       col = "red")

# Integral to 100 (slighlty arbitrary)
plot(sort(ints_simulations$int_to_100_norm),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to 100",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to 100 is ",
                   sum(ints_simulations$int_to_100_norm < pull(ints_simulations[1, "int_to_100_norm"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_100_norm),
       col = "red")
```

Looks like the miniscule rescaling of the kernel densities (see the first plot,
the integrals are only just above 1) makes little difference, as not rescaling
much. So don't bother with it.


### Plot integration results using spatstat integration function to compare to mine

```{r plotintsspatstat}
# Full integral
plot(sort(ints_simulations$full_spatstat),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Full integral",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for full integral is ",
                   sum(ints_simulations$full_spatstat < pull(ints_simulations[1, "full_spatstat"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(full_spatstat),
       col = "red")

# Integral to mean (which is the mode, so not overly surprising okay if not perfect)
plot(sort(ints_simulations$int_to_mean_spatstat),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to mean",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to mean is ",
                   sum(ints_simulations$int_to_mean_spatstat < pull(ints_simulations[1, "int_to_mean_spatstat"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_mean_spatstat),
       col = "red")

# Integral to 100 (slighlty arbitrary)
plot(sort(ints_simulations$int_to_100_spatstat),
     ints_simulations$sim_num,    # Not actually since sorted
     xlab = "Integral to 100",
     ylab = "Sorted simulations",
     main = paste0("Proportion less than theoretical for integral to 100 is ",
                   sum(ints_simulations$int_to_100_spatstat < pull(ints_simulations[1, "int_to_100_spatstat"]))/num_sims))

abline(v = filter(ints_simulations, sim_num == 0) %>%
         pull(int_to_100_spatstat),
       col = "red")
```

Looks like ..... TODO change this based on results: the miniscule rescaling of the kernel densities (see the first plot,
the integrals are only just above 1) makes little difference, as not rescaling
much. So don't bother with it.

Plot the integrals to 100 from spatstat versus Simpson's:
```{r spatsimp}
plot(ints_simulations$int_to_100,
     ints_simulations$int_to_100_spatstat,
     xlab = "Integral to 100 from Simpson's",
     ylab = "Integral to 100 from spatstat",
     main = paste0("Proportion Simpson's less spatstat is ",
                   sum(ints_simulations$int_to_100 < ints_simulations$int_to_100_spatstat)/num_sims))
abline(0, 1,
       col = "red")
```

So they're close, with maximum absolute difference of
`r max(abs(ints_simulations$int_to_100 - ints_simulations$int_to_100_spatstat))`,
which is a good vindication of my code (but doesn't solve the underlying hdi problem).

So we are good to switch integration methods. I think the lognormal may just be
the issue as it's skewed. Repeat this vignette for a normal distribution, in `simpsons-simulation-normal.Rmd`.
