---
title: "Simulation vignette"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("simulation-vignette.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

Moving what's needed to results-extra vignette, then can delete this. TODO

This investigation is what led to thinking
we should not use the density approach to calculated the HDI.

Doing some simulations to check some things, based on a simulated
values from a lognormal distribution, and comparing with known theoretical
HDI.

Issues 1 and 2 relate to our original write up. This code uses explicit
`density = TRUE` or `FALSE` arguments in `create_intervals()`, so will still
work if we change the default. Then use some of this for a simplified version to
be a real vignette if needed.

```{r setup}
load_all()      # TODO change to library(hdiAnalysis) when done
library(dplyr)
library(spatstat.geom)   # Needed for checking some integrals
# library(coda)            # Needed to test alternative code
```

A single set of simulated values are saved in the `sim` list object. Use those
first to explain, and then do repeated simulations further down.

```{r ints}
mean_log <- sim$mean_log
sd_log <- sim$sd_log
mean_log
sd_log
n <- length(sim$values)
n

ints_sim <- create_intervals(sim$values,
                             density = TRUE)
ints_sim$intervals %>% a()

# 'True' theoretical values, based on HDInterval calculation
hdi_true <- HDInterval::hdi(qlnorm, credMass = 0.95, meanlog = mean_log, sdlog = sd_log)

ints_true <- tibble::tibble(median = qlnorm(0.5, mean_log, sd_log),
                            eti_lower = qlnorm(0.025, mean_log, sd_log),
                            eti_upper = qlnorm(0.975, mean_log, sd_log),
                            hdi_lower = hdi_true["lower"],
                            hdi_upper = hdi_true["upper"]) %>%
  dplyr::mutate(width_eti = eti_upper - eti_lower,
                width_hdi = hdi_upper - hdi_lower,
                width_diff = width_eti - width_hdi,
                i_eti_lower = NA,
                y_eti_lower = dlnorm(eti_lower,
                                     mean_log,
                                     sd_log),
                i_eti_upper = NA,
                y_eti_upper = dlnorm(eti_upper,
                                     mean_log,
                                     sd_log),
                i_hdi_lower = NA,
                y_hdi_lower = dlnorm(hdi_lower,
                                     mean_log,
                                     sd_log),
                i_hdi_upper = NA,
                y_hdi_upper = dlnorm(hdi_upper,
                                     mean_log,
                                     sd_log),
                hdi_height = NA,
                integral_full = NA,
                integral_eti = plnorm(eti_upper,
                                     mean_log,
                                     sd_log) -
                               plnorm(eti_lower,
                                      mean_log,
                                      sd_log),
                integral_hdi = plnorm(hdi_upper,
                                     mean_log,
                                     sd_log) -
                               plnorm(hdi_lower,
                                      mean_log,
                                      sd_log),
                warning = NA,
                allow_hdi_zero = NA)

ints <- cbind(type = c("simulated",
                       "theoretical"),
              rbind(ints_sim$intervals,
                    ints_true))
ints

plot(ints_sim)
```

So `hdi_lower` seems to be lower for the simulated value than the theoretical
value (tried with multiple simulated data sets).

Check that theoretical calculations (which use `HDInterval::hdi.function()`) are correct.

First, the `create_intervals()` value:
```{r integrals}
# Calculate my simulated HDI width using plnorm
plnorm(ints_sim$intervals$hdi_upper,
       mean_log,
       sd_log) -
  plnorm(ints_sim$intervals$hdi_lower,
         mean_log,
         sd_log)
```

That is above 95%, different to the output from `create_intervals()` shown next because
that uses numerical.
```{r theorint}
select(ints, type, integral_hdi)
```

The theoretical value above is already calculated using `plnorm()`.


Maybe due to the trapezoid approximation. Best to write own
integral function, probably using Simpson's Rule. But the renormalising didn't
seem to be that important for this example, putting a browser in
`create_intervals()` so might not be the issue.

## Possible problems

1. `HDInterval::hdi.function()` is not giving the expected answer. Seems
   doubtful as checking the intergral separately seems to work. But it does
   minimise the width of the interval, rather than explicitly do the rising tide
   idea I think. Have dug into below, and all seems fine.

2. Numerical integration is not accurate. Renormalisation wasn't much of a
   factor within `create_intervals()` (based on a few examples). Make a Simpson's
   rule function. Done, but didn't help when using `density = TRUE`.

3. The kernel density estimate is causing issues. Turns out to be so, and
   repeated simulations below (related to what we had called Issue 1), show we should
   revert to using the default `HDInterval::hdi()`, not the density version.

## 1. Dig into `hdi.function()`

This is the plot of the quantile function.
```{r quantilefun}
probs <- seq(0, 1, 0.01)
plot(probs,
     qlnorm(probs, mean_log, sd_log),
     xlab = "Cumulative probability",
     ylab = "Data value")
abline(h = c(ints_true$hdi_lower,
             ints_true$hdi_upper),
       col = "red")
abline(v = c(plnorm(ints_true$hdi_lower,
                    mean_log,
                    sd_log),
             plnorm(ints_true$hdi_upper,
                    mean_log,
                    sd_log)),
       col = "red")

```

So `hdi.function()` optimises over all low-tail-probabilities to give the
shortest 95% interval (in the y-axis sense in the above figure) that starts
between 0 and 0.05. The inverse cumulative density function is used, basically
`qlnorm()`. And does it numerically but using functions so should be able to
hone in on the correct value. Lines added above should demonstrate the calculation.

So, optimising to give the shortest 95% interval. Given the HDI is the shortest
95% interval, this seems fine.

## 2. Numerical integration accuracy

Doesn't seem to the be the problem. Made a Simpson's rule function and compared
it to `spatstat::geom()` and gives the same results on kernels as that. See
`simpsons-simulate-normal.Rmd` vignette, which does find the numerical integrals
(for both methods) don't quite match the theoretical ones, which might just have
to live with.


## 3. Simulation experiment

Do multiple simulations. First with the same parameters for a lognormal, then
maybe need to do for multiple. Expect the multiple simulations the same parameters
should be fairly consistent.

Now also adding in issue 1 and 2 simulations, where issues are as described in
Extended Methods Appendix and `hdiAnalysis-vignette.Rmd` vignette (TODO that name
will likely change).

```{r simulations}
num_sims <- 1000
simulations <- list()
ints_simulations <- cbind(type = "theoretical",
                          ints_true)

ints_simulations_issue_1 <- ints_simulations
ints_simulations_issue_2 <- ints_simulations

for(i in 1:num_sims){
  set.seed(i)
  this_sim <- rlnorm(n,
                     mean_log,
                     sd_log)

  this_int_sim <- create_intervals(this_sim,
                                   density = TRUE)

  simulations[[i]] <- list("values" = this_sim,
                           "ints_sim" = this_int_sim)

  ints_simulations <- rbind(ints_simulations,
                            cbind(type = i,
                                  this_int_sim$intervals))

  # Issue 1
  # Use the methods as originally tried to explain Issue 1.
  this_int_sim_issue_1 <- create_intervals(this_sim,
                                           density = FALSE)
  # Not saving full results
  # simulations[[i]] <- list("values" = this_sim,
  #                         "ints_sim" = this_int_sim)

  ints_simulations_issue_1 <- rbind(ints_simulations_issue_1,
                            cbind(type = i,
                                  this_int_sim_issue_1$intervals))

  # Issue 2
  # Use the methods as originally tried to explain Issue 2.
  this_int_sim_issue_2 <- create_intervals(this_sim,
                                           density = TRUE,
                                           n = 512)
  # Not saving full results
  # simulations[[i]] <- list("values" = this_sim,
  #                         "ints_sim" = this_int_sim)

  ints_simulations_issue_2 <- rbind(ints_simulations_issue_2,
                            cbind(type = i,
                                  this_int_sim_issue_2$intervals))
}
ints_simulations <- tibble::as_tibble(ints_simulations)

ints_simulations <- dplyr::mutate(ints_simulations,
                                  hdi_int_plnorm = plnorm(hdi_upper,
                                                          mean_log,
                                                          sd_log) -
                                    plnorm(hdi_lower,
                                           mean_log,
                                           sd_log))

ints_simulations_issue_1 <- tibble::as_tibble(ints_simulations_issue_1)

ints_simulations_issue_1 <- dplyr::mutate(ints_simulations_issue_1,
                                          hdi_int_plnorm = plnorm(hdi_upper,
                                                                  mean_log,
                                                                  sd_log) -
                                            plnorm(hdi_lower,
                                                   mean_log,
                                                   sd_log))

ints_simulations_issue_2 <- tibble::as_tibble(ints_simulations_issue_2)

ints_simulations_issue_2 <- dplyr::mutate(ints_simulations_issue_2,
                                          hdi_int_plnorm = plnorm(hdi_upper,
                                                                  mean_log,
                                                                  sd_log) -
                                            plnorm(hdi_lower,
                                                   mean_log,
                                                   sd_log))
```

Now to plot the intervals:
```{r plotints}
plot(range(c(ints_simulations$hdi_lower, ints_simulations$hdi_upper)),
     c(1, nrow(ints_simulations)),
     type = "n",
     xlab = "95% interval",
     ylab = "Simulation row number")
lines(c(ints_simulations[1, "hdi_lower"],
        ints_simulations[1, "hdi_upper"]),
      c(1, 1),
      col = "red")
for(i in 1:num_sims){
  lines(c(ints_simulations[i+1, "hdi_lower"],
        ints_simulations[i+1, "hdi_upper"]),
      c(i+1, i+1),
      col = "black")
}
abline(v = c(ints_simulations[1, "hdi_lower"],
             ints_simulations[1, "hdi_upper"]),
       col = "red")
```

Show the histograms of lower and upper HDI values compared to true theoretical value:
```{r checklower}
hist(ints_simulations$hdi_lower)
abline(v = ints_simulations[1, "hdi_lower"], col = "red")

hist(ints_simulations$hdi_upper)
abline(v = ints_simulations[1, "hdi_upper"], col = "red")
```

So not too good. This is what led me to investigate further.

The integrals of the simulated intervals calculated by `create_intervals()` are
all good, with range of `r round(range(ints_simulations$integral_hdi[-1]), 5)`. But
this is expected as the intervals were defined based on the 95% integration of
the kernal density estimate.

If if we calculate them using `plnorm()` we get most over 95%, namely
`r sum(ints_simulations$hdi_int_plnorm[-1] > 0.95)` out of `r num_sims`.

So I think must be something wrong with the integrating or with the kernel
density estimates.

Reliance on the kernel density estimates was the issue, with integrals being
correct, but the kde adding extra error. Looks like what we called Issue 1 is
actually better:

## Issue 1 plots

Now to plot the intervals for issue 1
```{r plotints1}
plot(range(c(ints_simulations_issue_1$hdi_lower, ints_simulations_issue_1$hdi_upper)),
     c(1, nrow(ints_simulations_issue_1)),
     type = "n",
     xlab = "95% interval",
     ylab = "Simulation row number")
lines(c(ints_simulations_issue_1[1, "hdi_lower"],
        ints_simulations_issue_1[1, "hdi_upper"]),
      c(1, 1),
      col = "red")
for(i in 1:num_sims){
  lines(c(ints_simulations_issue_1[i+1, "hdi_lower"],
        ints_simulations_issue_1[i+1, "hdi_upper"]),
      c(i+1, i+1),
      col = "black")
}
abline(v = c(ints_simulations_issue_1[1, "hdi_lower"],
             ints_simulations_issue_1[1, "hdi_upper"]),
       col = "red")
```

Show the histograms of lower and upper HDI values compared to true theoretical
value:

```{r checklower1}
hist(ints_simulations_issue_1$hdi_lower)
abline(v = ints_simulations_issue_1[1, "hdi_lower"], col = "red")

hist(ints_simulations_issue_1$hdi_upper)
abline(v = ints_simulations_issue_1[1, "hdi_upper"], col = "red")
```

The integrals of the simulated intervals calculated by `create_intervals()` are
not perfect, with range of
`r round(range(ints_simulations_issue_1$integral_hdi[-1]), 5)`, but this is based on the
kde approximation.

If we calculate them using `plnorm()` we get the number being over 95% as
`r sum(ints_simulations_issue_1$hdi_int_plnorm[-1] > 0.95)` out of `r num_sims`.

## Issue 2 plots

Now to plot the intervals for issue 2
```{r plotints2}
plot(range(c(ints_simulations_issue_2$hdi_lower, ints_simulations_issue_2$hdi_upper)),
     c(1, nrow(ints_simulations_issue_2)),
     type = "n",
     xlab = "95% interval",
     ylab = "Simulation row number")
lines(c(ints_simulations_issue_2[1, "hdi_lower"],
        ints_simulations_issue_2[1, "hdi_upper"]),
      c(1, 1),
      col = "red")
for(i in 1:num_sims){
  lines(c(ints_simulations_issue_2[i+1, "hdi_lower"],
        ints_simulations_issue_2[i+1, "hdi_upper"]),
      c(i+1, i+1),
      col = "black")
}
abline(v = c(ints_simulations_issue_2[1, "hdi_lower"],
             ints_simulations_issue_2[1, "hdi_upper"]),
       col = "red")
```

Show the histograms of lower and upper HDI values compared to true theoretical
value:

```{r checklower2}
hist(ints_simulations_issue_2$hdi_lower)
abline(v = ints_simulations_issue_2[1, "hdi_lower"], col = "red")

hist(ints_simulations_issue_2$hdi_upper)
abline(v = ints_simulations_issue_2[1, "hdi_upper"], col = "red")
```

The integrals of the simulated intervals calculated by `create_intervals()` (so
using kde)
have a range of `r round(range(ints_simulations_issue_2$integral_hdi[-1]), 5)`.

If we calculate them using `plnorm()` we get the number over 0.95 being
`r sum(ints_simulations_issue_2$hdi_int_plnorm[-1] > 0.95)` out of `r num_sims`.
